### 2020-08-16 paper.tex
Rewrite RQ1 Answer

\subsection{Interpretation and Take Away}

@@

We were able to observe how all subjects modified the provided source code manually and augmented it with trace logs. All of them used the produced information in order to track down problems in the application. We could not observe the extraction to and reintegration from an external tool, which would have been yet another form of code modification.

For any subject, we could not identify any contradiction between the observed behavior and the answers given within the after-action survey.

All reviewed subjects exhibited the debugging behavior described in the initial hypothesis, which we can accept based on the collected evidence.

Interviewing professionals, evaluating the war stories, the hands-on experiences, and the results from the observational study allowed us to identify three main challenges, software engineers have to overcome when working with RxJS within an application: (i) They want to \emph{understand how and what data flows} through the graph described by the RP DSL, which helps to identify more straightforward problems most of the time. Though, knowing about the life cycle events of Observables was noted as crucial especially when working with more complex, higher-order compositions of Observables. (ii) Where traditional debugging tools provide support suited for an imperative programming style, they fall short when interacting with RP-based code which would require a \emph{deeper understanding of the RP runtimes internal graph-based model}. This circumstance makes it hard for engineers to retrace dataflows and computations. And finally, (iii) even though there are partial solutions available for the two previously described challenges, they usually require the software engineers either to \emph{manually modify source code or extracting parts of it} to an external utility. These additional efforts cost time and bring their own risks since the production code needs to be touched without adding actual value.

>>>>

\subsection{Insights}

@@

We were able to observe how all subjects predominantly used manual source code augmentation by adding trace logs. Only 50\% of them used traditional debugging utilities in order to inspect the programs state at runtime. All subjects used the new information gained to refine their hypothesis about underlying problems before they started a new iteration in the debugging process. We could not observe the extraction to and reintegration from an external tool.

Even though two subjects stated in the after-action survey to regularly use rxjs-spy for debugging RxJS programs, neither S1 nor S3 made use of this library during the experiment part of the study. That aside, we could not identify any contradiction between the observed behavior and the answers provided in the survey.

All reviewed subjects exhibited the debugging behavior described in our hypothesis, which we can accept based on the collected evidence.

Interviewing professionals, consolidating RxJS hands-on experiences from the war stories, and evaluating the results from our observational study showed us that software engineers use a variety of practices, tools and utilities to debug RP programs. Beside the habit of adding trace logs manually, we saw them evidently trying to answer their debugging hypotheses using traditional, imperative-focused debugger utilities. The later way of debugging was repeatedly commented as unsatisfying as these utilities cannot handle RP constructs and with this, cannot help to detect problems located within these at all. The former way, the introduction of manual log statements, was often described as ``the last resort'' when no other debugging technique helped before. This is because the analysis of log entries gets tedious as the number of those increases.

We heard further how engineers isolate specific Observables from bigger dataflows and how they inspect those in sandboxed environments and visualizers. This helps them understanding lifecycle and value emitting behaviors better and iterate faster in order to resolve problems.

We are able to postulate an answer to our first research question based on this collected information: The biggest challenge for software engineers debugging RxJS-based programs is knowing \emph{when} to apply \emph{what} tool in order to get the \emph{best possible} answer on their currently tested hypothesis. We further claim that, backed by the observation where two subjects knew about rxjs-spy but refrained from using it nonetheless, it is not enough to \emph{know} about the existence of a specific tool for a particular debugging situation. Such a tool should rather offer itself at the ``right'' moment so the engineer can benefit from its features. An example for such a behavior is the seamless integration of asynchronous call stacks in Google Chromes JavaScript debugger\cite{chromeasync}: Even though code was executed ``outside'' of the current call stack, the debugger shows these asynchronous stack frames as they were actually part of it.
